### 2020-01-28 paper/paper.tex
Luciana + me. Section 4

%\documentclass[a4paper, twocolumn]{article}
\documentclass[a4paper]{article}
\newcommand{\papertitle}{Potential of I/O-Aware Workflows in Climate and Weather}

\usepackage[a4paper, margin=2cm]{geometry}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage[colorlinks=true,urlcolor=red]{hyperref}
\usepackage{url}
\usepackage{float}
\usepackage{cleveref}
\usepackage{todonotes}
\usepackage{comment}
\usepackage{ulem}
\usepackage{array}
\usepackage{subcaption}

\newcommand{\jk}[1]{\todo[inline]{JK: #1}}
\newcommand{\lr}[1]{\todo[inline]{LR: #1}}

\usepackage{multicol}
\setlength{\columnsep}{1cm}

\usepackage{fancyhdr}
\fancyhead{}
\fancyfoot{}
\fancyhead[l]{\papertitle}
\fancyhead[r]{\includegraphics[height=1em]{esiwacelogo_type_grey_left_2c_ext}}
\fancyfoot[r]{\thepage}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{1pt}
\renewcommand{\footrulewidth}{1pt}

\usepackage{titling}
\pretitle{\begin{center}\Large\bfseries}
\posttitle{\end{center}\vskip 0.5em}

\graphicspath{{./assets/}}

\title{\papertitle}

\author{Julian M. Kunkel
  \textit{University of Reading}
    \and
  Luciana R. Pedro
  \textit{University of Reading}
}
\date{\today}

\begin{document}
\maketitle
\thispagestyle{fancy}

\section*{Abstract}
The efficient, convenient, and robust execution of data-driven workflows and enhanced data management are key for productivity in scientific computing.
Traditionally, in HPC, the concerns of storage and computing are separated and optimised independently from each other and from the needs of end-to-end user.
As climate and weather workflows become increasingly complex and blend beyond data centres while, at the same time, storage hierarchies become deeper, the community investigates means to re-organise storage access to utilise such environments fully.

The key contributions of this paper are:
1) we sketch the visions of an integrated data-driven approach and discuss the challenges and implications of this strategy;
2) we illustrate architecture and roadmap that allows the seamless integration into existing workflows.
The tools employed here to achieve this extended workflow are Cylc, XIOS, DDN IME, and ESDM.

%Basically, we believe workflows composed of data, computing, and communication-intensive tasks should drive the interfaces and hardware configurations to best support the programming models.
From the user perspective, we ultimately aim to provide an abstraction to data dependencies.
The proposed changes increase the opportunity of implementations for smarter scheduling of computing and storage in heterogeneous storage environments considering characteristics of workflows and data.
From the system perspective, the vision aims to derive an execution plan that utilises the available hardware and software infrastructure within a data centre.

\section{Introduction}

High-Performance Computing (HPC) harnesses the fastest hardware components to enable the execution of tightly coupled applications from science and industry.
Typical use-cases cover the numerical simulation of physical systems and the analysis of large-scale observational data.
In the domain of climate and weather, there is a considerable demand for the orchestration of ensembles of simulation models and the generation of data products.
A weather forecast service such as ECMWF generates .... \jk{Cylc team, please provide some facts}
Likewise, ensembles in climate can ...

Based on their needs, the HPC community has developed a software ecosystem that supports them to execute their large-scale workflows.
While the current advances correspond to a big leap forward, many processes still require experts.
For example, porting a workflow from one system to another requires to adjust runtime parameters of applications and to decide how data is managed.
Since performance is of crucial importance to large-scale workflows, careful attention must be paid to exploit the system characteristics of the target supercomputer.
To obtain the best performance, a specific supercomputer typically requires substantial changes to the workflow to tailor the workflow to the particular machine.
For instance, a data-driven workflow may benefit from using a heterogeneous set of computing and storage technology at the same time.

Knowing the capabilities, interfaces, and performance characteristics of individual components are mandatory to make the best use of them.
As the complexity of systems increases and alternative storage and computing technologies provide unique characteristics, it becomes increasingly difficult even for experts to manually optimise resource usage in workflows. In many cases, modifications are not performed because: 1) they are labour intense: any change to the workflow requires careful validation which may not pay off for small scale runs; 2) users are not aware of the potential of the complex system.

In this paper, we show that knowledge about Input/Output of workflow tasks and overall experimental design helps to
optimise the execution of climate and weather workflows and may increase the performance, throughput and cost-efficiency of the environments, providing an incentive to users and data-centres that cannot be neglected.

The paper is organised as follows:
In \Cref{sec:workflows}, we provide an overview of the specification of workflows in climate and weather and the involved software stack.

\jk{TODO}

\section{Workflows in Climate/Weather}
\label{sec:workflows}

In this section, we describe how the environment workflows are executed in the typical software stack.

\subsection{Data Center Infrastructure}

To satisfy the needs for different workflows, data centres are providing a infrastructure consisting of computing and storage devices with different characteristics which makes them more efficient for specific tasks.
Take, for example, the supercomputer Mistral at DKRZ that consists of 3,321 nodes.
It offers two types of compute nodes equipped with different CPUs and a range of GPU nodes.
Each node provides an SSD for local storage and DKRZ has additionally two shared Lustre file systems with different performance characteristics.
Individual users and projects are mapped to one file system explicitly and users can access it as work or scratch file system: While data is kept on the work file system indefinitely, its available space is limited by a quota.
The scratch file system allows to store more data, but data is automatically deleted after some time.

Future centres are expected to have even more heterogeneity. A variety of accelerators (GPU, TPU, FPGAs), active storage, in-memory, and in-network computing technologies provide storage and processing capabilities.
\Cref{fig:heterogeneous} shows such an environment with a focus on computation and storage.
Some of these technologies might be local to specific compute nodes or globally available.
Depending on the need, the storage characteristics range from predictable low-latency (in-memory storage, NVMe) to online storage (SSD, HDD), and cheap storage for long-term archival (tape).
\jk{Mention BBs, e.g., DDN IME}
Naturally, the variety of tasks executed by a single workflow may benefit from the different storage and computing infrastructure.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{system}
  \caption{Example of an heterogeneous HPC landscape}
  \label{fig:heterogeneous}
\end{figure}

\subsection{I/O Stack of a Parallel Application}

A typical I/O stack for a single parallel application like a climate model is shown in \Cref{fig:layers}.

\begin{minipage}{0.2\textwidth}
\begin{figure}[H]
  \includegraphics[width=\textwidth]{layers-xios}
  \caption{I/O Path for an MPI-Parallel Application}
  \label{fig:layers}
\end{figure}
\end{minipage}
\qquad
\begin{minipage}{0.7\textwidth}
In our example, we assume the application is parallelised for performance reasons and uses MPI to run the application.
It may use XIOS to gather data from individual fields which then uses NetCDF4 to store the data as a file.
XIOS provides a domain-level semantics and ...\jk{TODO}
Under the hood, NetCDF4 used the HDF5 API and file format.
Internally, HDF5 uses MPI again and its data types to specify the nature of the data stored.
Finally, data is stored on a parallel file system like Lustre, which, on the server-side, stores data in local file system as block devices on storage media such as SSDs and HDDs.

Different applications involved in a workflow may use a different I/O stack to store their outputs.
Naturally, the application that inputs previously generated data must use a compatible API to be able to read the specific data format.
In this figure, for example, XIOS is beneficial for parallel I/O and a process might directly rely on the NetCDF API to read previous data.
\end{minipage}

\subsection{Software Stack}

The relevant software stack involved in executing a workflow is depicted in \ref{fig:stages}.
As a preliminary, the user had specified a Cylc suite representing the workflow and provided scripts files for the individual tasks to execute; now the user enacts Cylc to start the workflow.
In the following, each stage of the execution is described further.

\begin{figure}[bt]
  \centering
  \includegraphics[scale=1.4]{stages}
  \caption{Software Stack and Stages of Execution}
  \label{fig:stages}
\end{figure}

\begin{enumerate}
  \item \textbf{Cylc}\ \ Cylc parses the suite file, generate dependencies and defines a schedule for the execution.
  Cylc monitors the progress of the workflow.
  Once a task can be executed (dependencies are fulfilled), Cylc generates a Slurm script on the fly with the required metadata for Slurm that will invoke the user-defined script for that task, and runs Slurm to queue the job.

  \item \textbf{Slurm}\ \ Schedules the execution of the job considering the scheduling policy of the data center.
  Once the job is ready to be dispatched, i.e., resources are available and the job priority is the highest, it is started on the supercomputer.

  \item \textbf{Job}\ \ The user-provided script is executed on a single node assigned to the job.
  The script has set environment variables containing information about the Cylc execution (e.g., task in the workflow) and Slurm job (e.g., number of nodes).

  \item \textbf{Script}\ \ The shell-script executes the sequence of commands including applications that are parallel programs.
  Typically, to create a filename, the script calls Cylc commands with a template provided by the user and feeds this information to the application.
  This ultimately defines the storage location to be used.
  A script typically consist of multiple fine-grained commands and may even call multiple parallel applications.

  \item \textbf{Application}\ \ The user-specified parallel application runs taking the filenames as defined in a configuration file or as arguments.
  The application may use NetCDF or XIOS to perform the I/O.
\end{enumerate}

\subsection{Cylc}

Cylc is in charge to execute and monitor workflows in which each step is submitted to the batch scheduler of a data centre.
 %, and the relevance of heterogeneous infrastructure to run workflows.
%We focus on high-level considerations to illustrate mandatory concepts.

Consider the Cylc suite configuration for a toy monthly cycling workflow in Figure \ref{fig:cylc}.
In this workflow, a warm-cycled atmospheric model (\textbf{model}) simulates the physics from a current state to predict the future after a timestep, for example, after six hours.
This process is repeated in the model to simulate month and years into the future.
Once the simulation of any month is computed, the data for this month becomes available and can be analysed.
A user has specified a detailed workflow in a Cylc suite file (\texttt{suite.rc}).
In this workflow, the \textbf{model} is followed by postprocessing (\textbf{post}), forecast verification (\textbf{ver}), and product generation (\textbf{prod}) tasks.
Also, consider, in addition, a task \textbf{check} that compares some verification metric against products from two cycles earlier.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.9\columnwidth]{cylc1.png}
  \caption{Example workflow\cite{8675433}}
  \label{fig:cylc}
\end{figure}

This suite presents the dependencies among the tasks, but it is missing the information about I/O and storage.
Decisions about storage must take into consideration the architecture in which the workflow will run.
The developers make the configuration about parallelism and where the data is stored in a script file that is executed for each specific task.

\subsection{Data Management}
\label{sec:datamanagement}

The scripts for the tasks in Cylc define how the data is stored on the available storage systems.
Consider that the machine has three file systems: a fast \textbf{scratch} file system on which data may reside only for a week, a slower \textbf{work} file system, and each node has a \textbf{local} file system.

What happens in many current workflows is that they use only \textbf{work} and \textbf{scratch}.
When a task is set to computing, the corresponding dataset is moved to \textbf{scratch}, processed, and the resulting datasets are transferred back to \textbf{work}.
If the \textbf{scratch} filesystem achieves its capacity, then the datasets move back to \textbf{work} and keep running the task until it is finished.
If the I/O planning is made a priori, this situation can be avoided.
In addition to the file systems described, many supercomputers also have available for the user \textbf{scratch2} file systems (sometimes local and global).
Therefore, in a real scenario, if the architecture of the machine is known, we can better utilize its resources.

\Cref{fig:lifecycle} shows three possible life cycles and the access pattern for a specific dataset.
As shown in \Cref{fig:lifecycle}(a), the dataset could be stored on the \textbf{local} storage first to avoid congestion on the \textbf{work} file system, then be migrated to \textbf{work} file system where subsequent tasks of the workflow may read it multiple times.
In the end, this dataset might be an intermediate product that can then be deleted.
Alternatively (\Cref{fig:lifecycle}(b)), it could be stored on the \textbf{scratch} file system immediately and accessed there.
However, that would require that the last read access happens before files on \textbf{scratch} are automatically removed.
The last scenario (\Cref{fig:lifecycle}(c)) presents the case where data is created on \textbf{work} by a task, but then it is copied to a \textbf{local} node to allow read accesses of subsequent tasks which might be beneficial for small random accesses.
However, it would require that the subsequent tasks are placed on the same machine where data is now stored.
The scientist responsible for the workflow must optimise the available resources intuitively.
These are just examples and currently the scientists perform such kind of optimisations manually.

\begin{figure}[tb]
  \centering
  \begin{subfigure}{.45\textwidth}
  \includegraphics[width=0.9\columnwidth]{lifecycle-1}
  \caption{Local and work}
  \end{subfigure}
  \begin{subfigure}{.45\textwidth}
  \includegraphics[width=0.9\columnwidth]{lifecycle-2}
  \caption{Scratch only}
  \end{subfigure}

  \begin{subfigure}{.45\textwidth}
  \includegraphics[width=0.9\columnwidth]{lifecycle-3}
  \caption{Local and work}
  \end{subfigure}

  \caption{Alternative life cycles for mapping a dataset to storage and the operations: \textbf{C}reate, \textbf{M}igrate, \textbf{R}ead, and \textbf{D}elete}
  \label{fig:lifecycle}
\end{figure}

\section{State of the Art}

Other tools for workflow handling, how is this information transported?

dataspaces + ADIOS + in-situ

\jk{TODO}

\section{Vision for I/O-Aware Workflows}

Applied scientists should not spend time understanding hardware characteristics, but using their time to develop their work, let the computer do the job the simplest way as possible, and just collect and analyse the results.
However, nowadays, to be able to run a job in a HPC environment efficiently, researchers have to have profound knowledge about their workflow, which is expected, but also about decisions regarding storage, communication, and computing.

We propose an approach to reduce the burden on researchers and, at the same time, optimise the decisions about jobs running in HPC systems.
Once we have an automated decision about where the job will run and how the storage will be managed, scientists can then run their workflows in any machine without further interactions and even without previous knowledge about the machine architecture.

Our vision requires two additional information: firstly, the user must augment the workflow description with information about IO requirements and explicitly annotate dependencies to data sets; secondly, the data center (or expert user) must provide a description of the storage architecture and characteristics for each storage system.

There are many approaches to implement the technology for this vision, it is apparent that changes need to be made to several software components to realize this vision.
In the next section, we will discuss the specific design for our transitional roadmap for climate and weather.

\subsection{Extended Workflow Description}

So far, Cylc allowed specifying tasks and the dependencies among them.
We aim to enrich this information with characteristics for input, and output, i.e., the datasets used.
An example workflow containing input data sets and (intermediate) products is illustrated in \Cref{fig:workflow}.
Nodes represent tasks or data; arrows indicate dependencies among between them.
In the example, Task\,1 needs two datasets to perform its work, it directly communicates with Task\,2 and produces Product\,1.
Most of the workflow can run automatically, except for the manual quality control of the products and the final data usage of Product\,3.
This last step represents the typical uncertainty of data reuse, i.e., it is unclear how Product\,3 will be used further.
In this approach, each task is annotated with the required inputs and the generated products.
Each data product is annotated with metadata such as data life cycle information, e.g., the value of data and how long should it be kept.

Indeed, by monitoring I/O accesses we can automatically draft such IO descriptions on behalf of the user to simplify the specification.
All we need is to run one cycle of the specific workflow, considering that a workflow cycle is executed many times in climate and weather, this is a cheap approach.
From the recorded IO accesses, we analyze the IO patterns and dependencies associated with it to generate the configuration file.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.4\columnwidth]{workflow}
  \caption{Example of a high-level workflow with tasks and data dependencies}
  \label{fig:workflow}
\end{figure}

\subsection{System Information}

The system information shall comprise of all available storage systems, the system topology, and the characteristics of each of the involved components.
This may include simplified or complex models for the component to approximate performance for specific I/O patterns allowing to predict performance behavior.
It is expected that the data center (or expert users) can create such a configuration file, for instance, by using vendor provided information or by executing benchmarks.

\subsection{Smarter IO Scheduling}

With this separation of concerns, we can abstract from the workflow what is essential and what a machine should optimise to ensure that the resources available are being used smartly.
In this paper, we focus on utilizing a heterogeneous storage environment instead of exhaustively discussing the potential that is unleashed with the right level of abstraction in place by researchers that exploit the available opportunities.
Every time parameters are manually tuned right now is room from improvement by using automatised decision making.
And Machine Learning has been proven effective in replacing human-beings and adapt the workflow dynamically in case of unexpected situations (the system is down, nodes are unresponsive, etc.).

Optimisations we discuss in this paper are related to the life cycle of datasets and the placement of such datasets into specific storage according to system performance characteristics and the workflow specification.
Firstly, the IO scheduler could make the placement decision for the individual data sets and manage their relocation during the life cycle as discussed in \Cref{sec:datamanagement}.

Consider the theoretical case to have just two tasks A and B that need to be repeated many times:
Datasets generated by Task\,\textbf{A} are used by Task\,\textbf{B} and hence both tasks must have access to the same data in the system.
Because there is one script responsible for setting the configuration, the decision in which directory to store data is somewhat fixed.
Such configuration was done manually and with restricted information about the storage and system architecture.
It would be interesting to explore storage options for the datasets and, e.g., have different cycles stored placed in different storage systems.
If the lifetime of such dataset is, for instance, two cycles, alternate the datasets into two scratch file systems is something that would be a simple job an I/O-aware scheduler.
However, manually, that implies at having at least two scripts\footnote{Complicated scripts would have allowed to change the storage type depending on the cycle but it is big burden to the user.} for each task with information about the different storage.

Figure \ref{fig:cycle-4} exemplifies a restriction of our example climate/weather workflow the way it was being generated.
Assume you have, for instance, four cycles of the model running with all the required post-processing steps.
Typically, the user has generated the scripts for each task by generating a filename using the same prefix that is then placed at the same storage type every time.
That happens because all datasets generated by Task \textbf{model} will be used by Task \textbf{post\,prod}, so they should stay on the same node.

The user-specific storage locations are a viable option to run a workflow.
The user might even have analysed various scenarios such as depicted in
\Cref{sec:datamanagement}.
However, by automating these tasks, we provide a smarter route taking into consideration the architecture available and the way the datasets are stored during the intermediate steps of the workflow.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.8\columnwidth]{cycle-4}
  \caption{Four stages workflow example}
  \label{fig:cycle-4}
\end{figure}

\subsection{Benefit}

Once we have the complete knowledge about how the workflow is expected to run, we can optimise steps that were before predefined by the user usually considering only options for feasibility.

\begin{itemize}
\item Abstraction: The user does not have to worry about the architecture that the workflow will run, removing the specialist from the decision-making process.

\item Optimization: The workflow will be optimised for the specific available system characteristics considering potentially even runtime characteristics.
This will increase performance and system utilization.

\item Performance-portability: The user has to specify the IO requirements only once for the tasks of the workflow.
The workflow can now be run on any system without user intervention.
Even more, if the system characteristics change, e.g.. it gets upgraded and an additional storage tier becomes available or if storage degrades, the I/O-aware workflow could automatically adapt and make use of this new environment.
\end{itemize}

\section{Specification of the Workflow from User Perspective}

From the user perspective, a general workflow can be described by three categories described in Table \ref{tab:workflow}.

\begin{table}[H]
\centering
\begin{tabular}{|m{2cm}|m{5.5cm}|m{3.5cm}|m{3.5cm}|}
\hline
& \parbox{5.5cm}{\centering \textbf{Input}} & \parbox{3.5cm}{\centering \textbf{Computing}} & \parbox{3.5cm}{\centering \textbf{Products}} \\ \hline \hline

Description & Information inserted in the workflow. & Tasks specified by the user. & Possible outcomes of the tasks. \\ \hline

\jk{?} File & \texttt{input.ds} & \texttt{tasks.ds} & \texttt{output.ds} \\ \hline

Composed of & Datasets/Scripts & Datasets/Scripts & Datasets \\ \hline

Information &

\begin{tabular}{m{5cm}}
\textbullet\ Dates for the experiments, how many times the workflow has to run, etc.\\
\textbullet\ Dependencies among the tasks.\\
\textbullet\ Information about the lifetime of the final outcome.\\
\textbullet\ Steps in which manual decisions have to be made.\\
\textbullet\ {{\color{cyan}{I/O information. (Where the datasets will be stored in each step of the workflow.)}}}\\
\end{tabular}

&

\begin{tabular}{m{3cm}}
\textbullet\ The data input.\\ \\
\textbullet\ The data output.\\ \\
\textbullet\ The data lifetime.\\ \\
\textbullet\ The dependencies regarding the other tasks.\\ \\
\textbullet\ {{\color{cyan}{I/O information.}}}
\end{tabular}

&

\begin{tabular}{m{3cm}}
\textbullet\ Datasets that will be used by other tasks.\\ \\
\textbullet\ Datasets that will generate the final output.\\
\end{tabular}

\\ \hline
\end{tabular}
\caption{Components of the workflow in the user perspective}
\label{tab:workflow}
\end{table}

For this work, Cylc is the essential tool regarding not only the execution of the workflow, but the exchange of information with the EWIOScheduler about I/O and storage.
Cylc receives, as predefined, information about the storage and dependencies of the datasets. The current information is used to construct and run the complete workflow, but it is not employed in optimising intermediate steps.
The configuration is already predefined, and it cannot change. For instance, if it is established that Task\,2 depends on Task\,1, then we have to wait for Task\,1 to be completed to start running Task\,2.
Thinking what about I/O, no decision is being made about an optimal way to store the data.
Consider, as an example, that the datasets from \texttt{product1.ds} (the final product of task 1) are divided into \texttt{product1-final.ds} and \texttt{product1-task2.ds}. Then, it is clear that \texttt{product1-task2.ds} will be required to run Task\,2, and, because of that, should be put in fast storage to quickly being retrieved, or even kept in local storage. The \texttt{product1-final.ds} dataset, however, can go directly to tape, if the lifetime of the final output has to be stored for a long time, but no access is required from the current workflow.
The bottom line here is: several options can be optimised if we know the whole workflow in advance.

\section{Design}
What changes now in terms of specification?

In this paper, we approach two ways to have the extra information from the user to be able to make decisions about storage and the processing of the datasets/tasks.

\begin{itemize}

\item The user will insert this information as an additional file to the Cylc workflow management. In fact, the user is already doing that, but he is doing that in just one specific way, to make the workflow possible. Here, we are thinking about simplifying the information the user is already providing. So, instead of giving the complete workflow to Cylc, we are expecting the user to provide straightforward information, like:

\begin{itemize}

\item Task\,1 usually requires ten hours running in a pc with standard configuration\footnote{Intel Core i5, 8 GB RAM, and 500 GB internal storage drive, for instance.}, requires 100 GB of storage that has to be persistent for no more than one week.

\item Task\,2 usually requires two hours running in a pc with standard configuration, requires 150 GB of storage that has to be persistent for no more than two weeks week.

\item Task\,3 usually requires five hours running in a pc with standard configuration, and it does not require extra storage other than the storage for the final results.

\end{itemize}

\item The second option to access this type of information from the user does not require any direct information at all. The idea here is to use the DDN IME Monitoring system and have an estimate for the parameters after one run of the complete workflow. This approach is beneficial to the user because s/he does not have to take any extra step that s/he is currently doing to run the workflow through Cylc. However, the price here is being paid with an additional run of the complete workflow, which can be demanding at both time and expenses (we are considering here the paid access to a supercomputer, for example). In the end, it is up to the user and the previous knowledge s/he has about the intermediate stages of the workflow.

\end{itemize}

Now, comes the question. Assuming the information is available, who will be responsible for making the decisions for optimising the storage and the processing tasks? The answer here is simple: the ESDM Scheduler will do it. The ESDM Scheduler can be used in different steps of the process. ESDM has already an interface with NetCDF. Assuming the user can provide extra information in advance, ESDM can interact in the process before the user give the data to the Cylc tool, using the information to construct an optimised workflow. Once Cylc builds the workflow, ESDM can interact again after the configuration of the scripts for both effective and efficient run of the parallel applications regarding the available storage and architecture.

\section{Extending Existing Workflows}

The focus of this paper is to exemplify the I/O optimisations that can be made if the datasets dependencies are established. This information, in fact, is presented in the workflows before the Cylc application, so nothing else will be required from the user. But to have the information explicitly, Cylc needs some changes to be able to:

\begin{enumerate}

\item deliver it together with the workflow, and

\item adapts the resulting workflow to consider the I/O modifications.

\end{enumerate}

Figure \ref{fig:cycle1} introduces one cycle of the suite in Figure \ref{fig:cylc} with a simplified notation. The arrows represent the tasks dependencies, but there is no information about the data that are generated. Tasks \textbf{verification} and \textbf{post\,proc} depend on Task \textbf{model}, but does Task \textbf{model} needs to be fully completed to satisfy the dependency? Of course, once Task \textbf{model} is completed, the dependent tasks can be executed, and that is the type of information Cylc has to create and reproduce the workflow.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.4\columnwidth]{cycle1}
  \caption{First stage workflow}
  \label{fig:cycle1}
\end{figure}

The scenario presented in Figure \ref{fig:cycle-io} shows the same cycle of Figure \ref{fig:cycle1}, but the extended information about datasets is introduced.
In this picture, we have the tasks dependencies, and we explicit the datasets dependencies regarding the tasks that generated them.
The way data moves through tasks is by creating datasets to pass on the information, including filename and storage destination. The datasets produced by every task are coded in the workflow scripts before Cylc evaluation, and it is just replicated by this tool. The tasks can create a different number of datasets, and these datasets may be necessary for the immediate next task in the current cycle, or tasks in different cycles.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{cycle-io}
  \caption{First stage workflow including datasets}
  \label{fig:cycle-io}
\end{figure}

Figure \ref{fig:cycle-io-dep} now introduces one example of dependencies among the datasets and tasks: datasets colours indicate which task will use their information and datasets in yellow represent information that will be used in a different cycle. The tricky part when we consider only tasks and not the datasets in the workflow is that it may lead to the false impression that the tasks only depend on the immediately previous one. Note, for instance, that one dataset produced by Task \textbf{model} is later used in Task \textbf{prod\,gen}. That is not an immediate conclusion when we analyse just the arrows for the dependencies. It is clear that Task \textbf{model} is a dependency for Task \textbf{post\,prod}, which is then a dependency for Task \textbf{prod\,gen}. Still, the crucial information here is that any dataset currently being used in the current cycle can also be used later. From the figure we can also see that Task \textbf{model} will generate one dataset for Task \textbf{post\,proc}, one dataset for Task \textbf{prod\,gen}, two datasets for Task \textbf{verification}, and three datasets for a later cycle.
The access to this information before the process starts can be used to optimise how these datasets will be stored during the workflow. This information today is static and manual, leaving opportunities for automatisation and optimisations open.
The idea here is to embrace the concept that tasks dependencies are really datasets dependencies. Once we can insert the datasets explicitly in the workflow, it is easier to see that many optimisations regarding storage can be made.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{cycle-io-dep}
  \caption{First stage workflow including datasets and dependencies}
  \label{fig:cycle-io-dep}
\end{figure}

\subsection{Modified Software Stack}

The relevant software stack involved in executing a workflow with our IO-aware scheduler is depicted in \Cref{fig:stages-io}.
In the following, we describe the modifications to the current execution (described in \Cref{fig:stages}) we propose in this vision paper.
Firstly, the end-user now has to provide an additional file that covers the IO information for each task; slight changes have to be made to the scripts that are executed, these will be discussed below.

\begin{figure}[tb]
  \centering
  \includegraphics[scale=1.4]{stages-io}
  \caption{Software Stack and Stages of Execution with our IO-Aware Scheduler}
  \label{fig:stages-io}
\end{figure}

\begin{enumerate}
  \item \textbf{Cylc}\ \ Cylc parses the suite file, generate dependencies and defines a schedule for the execution.
  Now, Cylc calls EIOS to identify potential to optimize the schedule.
  EIOS may have decided that subsequent jobs shall be placed on the same node or to reorder the execution of some jobs.
  Also it may have provided information for conducting data migration.

  \item \textbf{EIOS}\ \
  Reads information the workflow: the original suite file and the IO suit file.
  Acts depending on the stage of execution, i.e., we could be preparing execution in Cylc or we are already inside a script.
  It loads the \texttt{esdm.conf} file that contains the information about the available storage technology in the data center and their characteristics.
  Provides the information how to prioritize IO targets.

  A data management service may run as a service and purge or migrate data according to user life cycle specification and EIOS information.

  \item \textbf{Slurm}\ \ Schedules the execution of the job considering the scheduling policy of the data center.
  Cylc may now have added a constraint to place a job on the same nodes as a previous job which is handled by the modified Slurm.
  Also, if migrations have to be performed before or after a job is run, Slurm will administer them.

  \item \textbf{Job}\ \ The user-provided script is executed on a single node assigned to the job.

  The script has set environment variables containing information about the Cylc execution (e.g., task in the workflow) and Slurm job (e.g., number of nodes).

  \item \textbf{Script}\ \ The shell-script executes the sequence of commands.

  Typically, to create a filename, the script calls Cylc commands with a template provided by the user and feeds this information to the application.
  This ultimately defines the storage location to be used.
  A script typically consist of multiple fine-grained commands and may even call multiple parallel applications.

  The way the filenames are generated is now different; we will now use a replacement command that invokes a tool of EIOS to generate a pseudo filename.
  This filename will encode additional information for ESDM about how to prioritise data placement according to data access.
  Potentially, we also can generate an \texttt{esdm.conf} file to reflect the specific hardware we have available to run the specific application.

  \item \textbf{Application}\ \ The user-specified parallel application runs.
  The application may either use NetCDF with ESDM support or ESDM directly to access data sets.
  Hence, in \Cref{fig:layers}, the HDF5 layer is replaced with ESDM.
  ESDM loads the \texttt{esdm.conf} file that contains the information about the available storage backends to use and their configuration.
  Knowing the potential targets, ESDM schedules application-local I/O.
  ESDM extracts the long-term schedule information from the generated filenames and considers it during I/O scheduling to exploit the data locality between tasks.

\end{enumerate}

\section{Conclusions}
\label{sec:conclusions}

\begin{comment}

and computer-aided Research Development and Engineering (RD\&E)

and also on small-scale systems.

For example, Big data tools integrate computing and storage capabilities into a holistic solution demonstrating the benefit of tight integrating for data-driven workflows.

%-- the quality of service features available in hardware is rarely used.

Cloud definition of desired state and contracts.

Indeed many research prototypes address subproblems
* But not all aspects together
* Competing approaches; the standardisation we describe here does not compete!

In this paper, we purposely do not mention too many of the ongoing RD\&E in storage and computing.

To wrap up, NGI provides a high-level view on data and processing on heterogeneous platforms while shielding the user from execution details.

Open ...

%Kubernetes

\begin{description}

\item{NFS} Relatively slow, persistent file system - Nightly incremental backups

\item{Lustre} Large, fast distributed scratch file system - No backups

\url{https://info.hpc.sussex.ac.uk/hpc-guide/storage.html}

\end{description}

%, and data products are generated, which are processes that are executed outside of the first simulation.

\end{comment}

\section{Challenges and Limitations}
\label{sec:challenges}
The current HPC environment faces several challenges and limitations concerning the execution of data-driven workflows.

\begin{description}
\item[Explicit Workflows]

Typically, as part of shell scripts, users explicitly encode data processing workflows as a sequence of steps consisting of input, computing, and \sout{{\color{blue}{output}}} {\color{cyan}{products}}.
The computing-intense parts ({\color{blue}{jobs}} {\color{cyan}{tasks?}}) and their dependencies are administrated by a cluster resource manager that dispatches them.
These loosely coupled workflows require that intermediate results are stored on persistent storage to be used by the next task.
Data transfer, however, needs a dominant fraction of energy compared to computation favouring offloading of computation to storage.

\item[Suboptimal Storage Tiering]

Organising the data placement on storage tiers is typically performed manually by the developer/user or via policies, leading to suboptimal optimisations. On top of that, it is a tedious and error-prone task.

Manual tiering requires the user or application to control the data placement, i.e., storing data (typically in the form of \sout{{\color{blue}{files}}} {\color{cyan}{datasets}}) on a particular storage system and, usually, moving data between storage by using scripts.
One limitation here is that decisions about how data structures are mapped and packaged into \sout{{\color{blue}{files}}} {\color{cyan}{datasets}}) are made once by the producing application, and cannot be changed without manual intervention by a downstream application.

A policy system (e.g. burst buffer) aims to simplify the data movement for the user, but typically migrates objects in the coarse granularity of files.
However, the semantical information that can be used by a policy system to make decisions is limited, e.g., data location, file extension, age of the file, etc.

Any hard-coded decision has significant implications for the performance of both applications and systems because the optimal choices about file size and content may be very different for the data production and any subsequent data analysis phase.
In any case, data migration needs resources from the source {\color{cyan}{source? the machine?}} and the target storage system to copy data.
Thus, to avoid extra manual work and decisions, it is preferable to direct data to locations where it can be stored and directly used by subsequent workflows -- at best, coupling \sout{{\color{blue}{the data source producing data with a data consumer}}} {\color{cyan}{the data source with the data consumer}}) to prevent storing intermediate products on persistent storage.

\item[Task and Data Equality]

All tasks and data products are handled identically by a HPC system, meaning there is no differentiation between individual jobs and data products.
This lack of hierarchy is problematic, as different use cases may not only differ in priority and deadline, but the computing speed usually depends on the critical path of the execution.

This problem is even more severe for data products: a storage system provides a single level of performance and resilience \sout{{\color{blue}{and performance}}} for stored files.
\sout{{\color{blue}{files}}} {\color{cyan}{datasets}})
Theoretically, different fault-tolerance and performance policies can be realised with varying storage systems and using management policies.
However, these policies would be coarse-grained and would not account for the priority of the workflow and, most importantly, the value of data.
The importance of data depends on aspects like costs to reproduce the data (can it be reproduced easily), the type of the experiment (test run, production run) and runtime constraints for the overall and potential crucial workflow.
While value and priority should influence fault-tolerance strategies and imply the quality of service for performance and availability, in practice, this is not the case as the storage just sees blocks and files. {\color{cyan}{I don't think we will be dealing with that in the next 10 years!}}

\item[Manual Optimisation]

Utilising the computing and storage characteristics of a heterogeneous environment as depicted in \Cref{fig:heterogeneous} is challenging.
Even within a single data centre, utilising homogeneous storage and computing infrastructure efficiently is complex for experts and needs manual definition.
The efficient management of data and computing capabilities in a heterogeneous environment and
how the execution of individual tasks from workflows may benefit from alternative hardware architectures and infrastructures
are still unresolved questions.

The optimisation of a single parallel job already is quite challenging.
To make optimal use of available features, programmers and users are forced to think about technical aspects and terms, e.g., setting low-level MPI hints for file striping, while they understand their workflow.
Similarly, efficient data placement is complex. But a specific workflow combined with manual optimisation is the root-cause of subsequent problems, and it must be avoided and replaced with automated and optimised decisions.

Manual and hard-coded workflows cannot handle changes in the environment and are error-prone.
Therefore, users must be able to express their workflows in an abstract fashion that allows the system to generate (near-)optimal execution plans and monitor their execution.

\end{description}

\section{Vision}

Our vision for the problems described in the first sections is a new approach for data management and data-driven computation.

\begin{description}
\item[Workflow Specification]

The user defines a workflow describing the intentions of data manipulation and the data lifecycle, including manual (user) involvement in the data processing.
The definition of constraints, for instance, deadlines for the generation of the final data products (insight), can also be included.
Workflows can be defined using arbitrary tools, including Python.
Note that a user will not provide an accurate mapping to system or infrastructure, but expects that the system derives an optimised execution plan that covers heterogeneous infrastructures.
It does not matter where data is processed and how and where information is stored, but only that the desired result is computed with all constraints met.
Workflows and data lineage of data products are recorded and can be inspected by the user upon demand, allowing reproduction and validation of the correct execution of experiments.

\item[Smart Scheduling]

Supplied with user workflows and their characteristics, the system derives execution plans for the computation and make data placement decisions that aim to utilise the available hardware resources efficiently.
As the API is in charge of both tasks, it may decide to trade computation and storage capacity considering characteristics as costs and energy consumption, for example. Additionally, it can enable automatic recomputation of intermediate states utilising virtualisation and container technologies.
Similarly, the scheduler may decide to couple subsequent steps of a workflow directly without storing intermediate data products on persistent storage.
Data might be replicated but enables the system to rerun parts of the workflow in case of a data loss.

Automatic scheduling is key to efficient processing and will allow constant improvement without forcing the user to adjust their workflows.
In most cases, individual resources are not used exclusively by a single task, but QoS methods ensure that constraints of workflows are met.
Note that this is a significant change for HPC environments, as computing resources are typically dedicated to specific jobs.

\end{description}

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{workflow1-v2}
  \caption{Workflow 1.}
  \label{fig:work1}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{workflow2-v2}
  \caption{Workflow 2.}
  \label{fig:work2}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{workflow-all}
  \caption{Workflow All.}
  \label{fig:work2}
\end{figure}

\begin{comment}

\begin{table}[H]
  \centering
  \begin{tabular}{cc}
  \includegraphics[scale=0.6]{workflow1-v2} &
  \includegraphics[scale=0.6]{workflow2-v2}
  \end{tabular}
  \caption{Workflows 1 and 2.}
  \label{fig:work1}
\end{table}

\end{comment}

\begin{comment}

\begin{description}

\item[Input] The information inserted in the workflow.

\item[Tasks] The tasks specified by the user.

\item[Products] Any of the possible outcomes of the tasks.

For simplicity, we will call these datasets as \texttt{product-workflow.ds}. In this specific example, it is composed by the datasets \texttt{product1.ds}, \texttt{product2.ds}, and \texttt{product3.ds}.

\end{description}

\end{comment}

\begin{comment}

Now.

\begin{enumerate}

  \item \textbf{Suite}\ \ The user runs Cylc to start a suite. The file \texttt{suite.rc} is created.

  \item \textbf{Cylc}\ \ Cylc parses the suite file, generate dependencies and defines a schedule for the execution.

  \item \textbf{Slurm}\ \ Once a task can be executed (dependencies are fulfiled), Cylc generates on the fly a Slurm script with the required metadata for Slurm, invokes the user-defined script for that task, and runs Slurm to queue the process.

  \item \textbf{Task}\ \ Once the task is ready to be dispatched by Slurm, it is started on the supercomputer.

  \item \textbf{Script}\ \ The shell script executes the users-provided instructions allowing Cylc to generate its internal variables. To create the filenames, Cylc reads templates provided by the user that defines the storage location to be used.

  \item \textbf{Application}\ \ The user-specified application runs taking the filenames as instructed. The application may use NetCDF or XIOS to perform the I/O.

\end{enumerate}

Our Plan.

\end{comment}

\bibliographystyle{alpha}
\bibliography{paper}

\end{document}